---
title: "Formative Assessment 7"
author: "Vera Aguila"
date: "2024-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Independent Samples T-Test: Invisibility Cloak

The dataset "Invisibility Cloak" explores the number of mischievous acts committed by two groups: those with an invisibility cloak and those without. We will perform an independent samples t-test to determine if having an invisibility cloak significantly affects mischievous behavior.


The dataset contains the following variables:

* Participant: Participant identification number.
* Cloak: Experimental group (0 = without a cloak, 1 = with a cloak).
* Mischief: The number of mischievous acts committed.


```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
```

```{r}
data <- data.frame(
  Participant = 1:24,
  Cloak = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
  Mischief = c(3, 2, 4, 5, 3, 2, 4, 5, 6, 3, 2, 4, 5, 6, 8, 7, 9, 6, 7, 8, 9, 10, 6, 7)
)

head(data)
```


## Assumption 1: The Dependent Variable is Continuous
The dependent variable, Mischief, is measured at a continuous level.


<br>

## Assumption 2: The Independent Variable has Two Categorical Groups
The independent variable Cloak consists of two groups:

1. Without a cloak
2. With a cloak


```{r}
table(data$Cloak)
```


<br>

## Assumption 3: Independence of Observations
Each participant is assigned to only one group. This assumption is met since no participant appears in both groups.


```{r}
# Boxplot to check for outliers
library(ggplot2)

ggplot(data, aes(x = factor(Cloak), y = Mischief)) +
  geom_boxplot() +
  labs(title = "Boxplot of Mischief by Cloak Group", x = "Cloak", y = "Mischief")

```


<br>

## Assumption 4: No Significant Outliers
We will check for outliers in the Mischief variable using boxplots.

Based on the visual inspection, we can determine if there are any extreme outliers.


## Assumption 5: Normality of the Dependent Variable
We will use the Shapiro-Wilk test to check if the Mischief variable is normally distributed for each group.

```{r}
# Shapiro-Wilk test for normality
shapiro.test(data$Mischief[data$Cloak == 0])  # Without Cloak
shapiro.test(data$Mischief[data$Cloak == 1])  # With Cloak

```

If p > 0.05, the data is normally distributed.


## Independent Samples T-Test
Finally, we will compute the independent samples t-test to compare the means between the two groups.

```{r}
t.test(Mischief ~ Cloak, data = data, var.equal = TRUE)  # Assuming equal variances
```

## Interpretation
If p < 0.05, we reject the null hypothesis and conclude that there is a significant difference in the number of mischievous acts committed between the two groups.



